Supplementary Material for the article \textit{Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures}

### 2017-10-13 experiments.tex
typos

When predicting $P(d_{test}|D_{train},S,\theta)$,
with structures $S$ learned by the BDeu score, we used the
Bayesian predictive parameter values (BPP) $\theta_{ijk} \propto
N_{ijk}+\frac{1}{r_iq_i}$.  In the spirit of keeping the scores
hyperparameter-free, for structures learned by the other model
selection criteria, we used the sequential predictive NML (sNML)
parametrization $\theta_{ijk}\propto e(n_{ijk})(n_{ijk}+1)$, where
$e(n)=(\frac{n+1}{n})^n$ as suggested in~\cite{Riss07b}.

@@

The graphs for different sample sizes for both predictive accuracy and
the number of parameters can be found in the Supplementary Material.

>>>>

When predicting $P(d_{test}|D_{train},S,\theta)$,
with structures $S$ learned by the BDeu score, we used the
Bayesian predictive parameter values (BPP) $\theta_{ijk} \propto
N_{ijk}+\frac{1}{r_iq_i}$.  In the spirit of keeping the scores
hyperparameter-free, for structures learned by the other model
selection criteria, we used the sequential predictive NML (sNML)
parametrization $\theta_{ijk}\propto e(N_{ijk})(N_{ijk}+1)$, where
$e(n)=(\frac{n+1}{n})^n$ as suggested in~\cite{Riss07b}.

@@

The graphs for different sample sizes for both predictive accuracy and
the number of parameters can be found in Appendix D in the Supplementary Material.

### 2017-10-13 qnml.tex
small typos
typos

We will next state this as a
theorem and sketch a proof for it. A more detailed proof appears in Appendix A the Supplementary Material.
\begin{theorem}
  \label{thm:scoreqv}
  Let $G$ and $G'$ be two Bayesian network structures that differ only
  by a single covered arc reversal, i.e., the arc from $A$ to $B$ in $G$
  has been reversed in $G'$ to point from $B$ to $A$, then
  $$s^{qNML}(D;G)=s^{qNML}(D;G').$$
\end{theorem}
\begin{proof}
  Now the scores for structures can be decomposed as
  $s^{qNML}(D;G)=\sum_{i=1}^{n}s_i^{qNML}(D;G)$ and
  $s^{qNML}(D;G')=\sum_{i=1}^{n}s_i^{qNML}(D;G')$.  Since only the
  terms corresponding to the variables $A$ and $B$ in these sums are
  different, it is enough to show that the sum of these two terms are
  equal for $G$ and $G'$. Since we can assume the data to be fixed we
  lighten up the notation and write
  $P^1_{NML}(i,G_i) := P^1_{NML}(D_{i,G_i};G)$ and
  $P^1_{NML}(G_i)   := P^1_{NML}(D_{G_i};G)$.
  \begin{eqnarray}
    \lefteqn{s_A^{qNML}(D;G)+s_B^{qNML}(D;G)} \nonumber\\
    && =\log\frac{P^1_{NML}(A,G_{A})}{P^1_{NML}(G_{A})}
            \frac{P^1_{NML}(B,G_{B})}{P^1_{NML}(G_{B})}\nonumber\\
    && =\log 1\cdot\frac{P^1_{NML}(B,G_{B})}{P^1_{NML}(G_{A})}\nonumber\\
    && =\log \frac{P^1_{NML}(B,G'_{B})}{P^1_{NML}(G'_{A})}
             \frac{P^1_{NML}(A,G'_{A})}{P^1_{NML}(G'_{B})}\nonumber\\
 && =s_A^{qNML}(D;G')+s_B^{qNML}(D;G'),\nonumber
\end{eqnarray}
  using the equations $\{A\}\cup G_A = G_B$, $\{B\}\cup G'_B = G'_A$,
  $\{B\}\cup G_B = \{A\} \cup G'_A$, and $G_A = G'_B$ which follow
  easily from the definition of covered arcs.
\end{proof}

\subsection{qNML Is Consistent}

@@

\subsection{qNML Is Regular}

@@

By analyzing the penalty term of the qNML scoring function, one can prove the following statement:
\begin{theorem}
qNML score is regular.
\end{theorem}
\begin{proof}
The proof is given in Appendix B in the Supplementary Material.
\end{proof}

>>>>

We will next state this as a
theorem and sketch a proof for it. A more detailed proof appears in Appendix A in the Supplementary Material.
\begin{theorem}
  \label{thm:scoreqv}
  Let $G$ and $G'$ be two Bayesian network structures that differ only
  by a single covered arc reversal, i.e., the arc from $A$ to $B$ in $G$
  has been reversed in $G'$ to point from $B$ to $A$, then
  $$s^{qNML}(D;G)=s^{qNML}(D;G').$$
\end{theorem}
\begin{proof}
  Now the scores for structures can be decomposed as
  $s^{qNML}(D;G)=\sum_{i=1}^{n}s_i^{qNML}(D;G)$ and
  $s^{qNML}(D;G')=\sum_{i=1}^{n}s_i^{qNML}(D;G')$.  Since only the
  terms corresponding to the variables $A$ and $B$ in these sums are
  different, it is enough to show that the sum of these two terms are
  equal for $G$ and $G'$. Since we can assume the data to be fixed we
  lighten up the notation and write
  $P^1_{NML}(i,G_i) := P^1_{NML}(D_{i,G_i};G)$ and
  $P^1_{NML}(G_i)   := P^1_{NML}(D_{G_i};G)$.
  \begin{eqnarray}
    \lefteqn{s_A^{qNML}(D;G)+s_B^{qNML}(D;G)} \nonumber\\
    && =\log\frac{P^1_{NML}(A,G_{A})}{P^1_{NML}(G_{A})}
            \frac{P^1_{NML}(B,G_{B})}{P^1_{NML}(G_{B})}\nonumber\\
    && =\log 1\cdot\frac{P^1_{NML}(B,G_{B})}{P^1_{NML}(G_{A})}\nonumber\\
    && =\log \frac{P^1_{NML}(B,G'_{B})}{P^1_{NML}(G'_{A})}
             \frac{P^1_{NML}(A,G'_{A})}{P^1_{NML}(G'_{B})}\nonumber\\
 && =s_A^{qNML}(D;G')+s_B^{qNML}(D;G'),\nonumber
\end{eqnarray}
  using the equations $\{A\}\cup G_A = G_B$, $\{B\}\cup G'_B = G'_A$,
  $\{B\}\cup G_B = \{A\} \cup G'_A$, and $G_A = G'_B$ which follow
  easily from the definition of covered arcs.
\end{proof}

\subsection{qNML is Consistent}

@@

\subsection{qNML is Regular}\label{sec:regularity}

@@

By analyzing the penalty term of the qNML scoring function, one can prove the following statement:
\begin{theorem}
qNML score is regular.
\end{theorem}
\begin{proof}
The proof is given in Appendix B in the Supplementary Material.
\end{proof}
As fNML criterion differs from qNML only by how the penalty term is defined, we obtain the following result with little extra work:
\begin{theorem}
fNML score is regular.
\end{theorem}
\begin{proof}
The proof is given in Appendix B in Supplementary Material.
\end{proof}
